Experiments with Synthetic Data

### 2022-07-31 dev/notebooks/housing.qmd
wuoohoo

---
title: Experiments with Housing Data
jupyter: julia-1.7
---

@@

```{julia}
include("dev/utils.jl")
using AlgorithmicRecourseDynamics
using CounterfactualExplanations, Flux, Plots, PlotThemes, Random, LaplaceRedux, LinearAlgebra
theme(:wong)
output_path = output_dir("real_world")
www_path = www_dir("real_world")
```

>>>>

---
title: Preprocessing Housing Data
jupyter: julia-1.7
---

@@

```{julia}
include("dev/utils.jl")
using AlgorithmicRecourseDynamics
using CounterfactualExplanations, Flux, Plots, PlotThemes, Random, LaplaceRedux, LinearAlgebra
theme(:wong)
output_path = output_dir("real_world")
www_path = www_dir("real_world")
data_path = data_dir("real_world")
```

```{julia}
using MLDatasets, Statistics, StatsBase
X = Matrix(BostonHousing().features)
dt = fit(ZScoreTransform, X, dims=1)
StatsBase.transform!(dt, X)
using DataFrames
df = DataFrame(X,:auto)
# Target:
y = BostonHousing().targets.MEDV
y = Float64.(y .>= median(y)); # binary target
df.target .= y
using CSV
CSV.write(joinpath(data_path, "housing.csv"), df)
```

### 2022-07-31 dev/notebooks/real_world.qmd
wuoohoo

---
title: Experiments with Real-World Data
jupyter: julia-1.7
---

@@

## Models

```{julia}
# Models:
model = Models.build_mlp()
λ = 0.1
model_laplace = Laplace(model, λ=λ)
M = FluxModel(model)
M_laplace = LaplaceReduxModel(model_laplace)
models = Dict(:laplace=>M_laplace, :mlp=>M)
```

## Generators

>>>>

---
title: Experiments with Credit Data
jupyter: julia-1.7
---

@@

```{julia}
catalogue = AlgorithmicRecourseDynamics.Data.load_real_world()
choices = [:give_me_some_credit_balanced, :housing]
data_sets = filter(p -> p[1] in choices, catalogue)
```

```{julia}
using CounterfactualExplanations.DataPreprocessing: unpack
bs = 50
function data_loader(data::CounterfactualData)
    X, y = unpack(data)
    data = Flux.DataLoader((X,y),batchsize=bs)
    return data
end
model_params = (batch_norm=false,n_hidden=128,dropout=true,p_dropout=0.1)
```

```{julia}
models = [:FluxModel, :FluxEnsemble]
generators = Dict(
    :Greedy=>GreedyGenerator(), 
    :Generic=>GenericGenerator(),
    :REVISE=>REVISEGenerator(),
    :DICE=>DiCEGenerator(),
    :Gravitational=>GravitationalGenerator(),
)
```

```{julia}
experiments = set_up_experiments(data_sets,models,generators; pre_train_models=100, model_params=model_params, data_loader=data_loader)
```

```{julia}
# #| eval: false
# using AlgorithmicRecourseDynamics.Models: model_evaluation
# plts = []
# for (exp_name, exp_) in experiments
#     for (M_name, M) in exp_.models
#         score = round(model_evaluation(M, exp_.test_data),digits=2)
#         plt = plot(M, exp_.test_data; title="$exp_name;\n $M_name ($score)", dim_red=:umap)
#         # # Errors:
#         # ids = findall(vec(round.(probs(M, exp_.test_data.X)) .!= exp_.test_data.y))
#         # x_wrongly_labelled = exp_.test_data.X[:,ids]
#         # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
#         plts = vcat(plts..., plt)
#     end
# end
# plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
# savefig(plt, joinpath(www_path,"models_test_before.png"))
```

```{julia}
# #| eval: false
# using AlgorithmicRecourseDynamics.Models: model_evaluation
# plts = []
# for (exp_name, exp_) in experiments
#     for (M_name, M) in exp_.models
#         score = round(model_evaluation(M, exp_.train_data),digits=2)
#         plt = plot(M, exp_.train_data; title="$exp_name;\n $M_name ($score)", dim_red=:umap)
#         # # Errors:
#         # ids = findall(vec(round.(probs(M, exp_.train_data.X)) .!= exp_.train_data.y))
#         # x_wrongly_labelled = exp_.train_data.X[:,ids]
#         # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
#         plts = vcat(plts..., plt)
#     end
# end
# plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
# savefig(plt, joinpath(www_path,"models_train_before.png"))
```

## Simple convergence

```{julia}
n_evals = 1
n_rounds = 10
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 2
n_bootstrap = 1
using Serialization
results = run_experiments(
    experiments;save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds, n_folds=n_folds, n_bootstrap=n_bootstrap
)
Serialization.serialize(joinpath(output_path,"results.jls"),results)
```

```{julia}
# using AlgorithmicRecourseDynamics.Models: model_evaluation
# plot_dict = Dict(key => Dict() for (key,val) in results)
# fold = 1
# for (name, res) in results
#     exp_ = res.experiment
#     plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
#     rec_sys = exp_.recourse_systems[fold]
#     sys_ids = collect(exp_.system_identifiers)
#     M = length(rec_sys)
#     for m in 1:M
#         model_name, generator_name = sys_ids[m]
#         M = rec_sys[m].model
#         score = round(model_evaluation(M, exp_.test_data),digits=2)
#         plt = plot(M, exp_.test_data; title="$name;\n $model_name ($score)", dim_red=:umap)
#         # # Errors:
#         # ids = findall(vec(round.(probs(M, exp_.test_data.X)) .!= exp_.test_data.y))
#         # x_wrongly_labelled = exp_.test_data.X[:,ids]
#         # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
#         plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
#     end
# end
# plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
# for (name, plts) in plot_dict
#     plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
#     savefig(plt, joinpath(www_path,"models_test_after_$name.png"))
# end
```

```{julia}
# using AlgorithmicRecourseDynamics.Models: model_evaluation
# plot_dict = Dict(key => Dict() for (key,val) in results)
# fold = 1
# for (name, res) in results
#     exp_ = res.experiment
#     plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
#     rec_sys = exp_.recourse_systems[fold]
#     sys_ids = collect(exp_.system_identifiers)
#     M = length(rec_sys)
#     for m in 1:M
#         model_name, generator_name = sys_ids[m]
#         M = rec_sys[m].model
#         data = rec_sys[m].data
#         score = round(model_evaluation(M, data),digits=2)
#         plt = plot(M, data; title="$name;\n $model_name ($score)", dim_red=:umap)
#         # # Errors:
#         # ids = findall(vec(round.(probs(M, data.X)) .!= data.y))
#         # x_wrongly_labelled = data.X[:,ids]
#         # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
#         plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
#     end
# end
# plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
# for (name, plts) in plot_dict
#     plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
#     savefig(plt, joinpath(www_path,"models_train_after_$name.png"))
# end
```

```{julia}
using Serialization
results = Serialization.deserialize(joinpath(output_path,"results.jls"))
```

## Strict convergence

```{julia}
n_evals = 1
n_rounds = 10
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 2
n_bootstrap = 1
using Serialization
results = run_experiments(
    experiments;
    save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds, n_folds=n_folds, n_bootstrap=n_bootstrap,
    convergence=:strict, save_name_suffix="strict"
)
Serialization.serialize(joinpath(output_path,"results_strict.jls"),results)
```

```{julia}
using AlgorithmicRecourseDynamics.Models: model_evaluation
plot_dict = Dict(key => Dict() for (key,val) in results)
fold = 1
for (name, res) in results
    exp_ = res.experiment
    plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
    rec_sys = exp_.recourse_systems[fold]
    sys_ids = collect(exp_.system_identifiers)
    M = length(rec_sys)
    for m in 1:M
        model_name, generator_name = sys_ids[m]
        M = rec_sys[m].model
        score = round(model_evaluation(M, exp_.test_data),digits=2)
        plt = plot(M, exp_.test_data; title="$name;\n $model_name ($score)", dim_red=:umap)
        # # Errors:
        # ids = findall(vec(round.(probs(M, exp_.test_data.X)) .!= exp_.test_data.y))
        # x_wrongly_labelled = exp_.test_data.X[:,ids]
        # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
        plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
    end
end
plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
for (name, plts) in plot_dict
    plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
    savefig(plt, joinpath(www_path,"models_test_after_$(name)_strict.png"))
end
```

```{julia}
using AlgorithmicRecourseDynamics.Models: model_evaluation
plot_dict = Dict(key => Dict() for (key,val) in results)
fold = 1
for (name, res) in results
    exp_ = res.experiment
    plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
    rec_sys = exp_.recourse_systems[fold]
    sys_ids = collect(exp_.system_identifiers)
    M = length(rec_sys)
    for m in 1:M
        model_name, generator_name = sys_ids[m]
        M = rec_sys[m].model
        data = rec_sys[m].data
        score = round(model_evaluation(M, data),digits=2)
        plt = plot(M, data; title="$name;\n $model_name ($score)", dim_red=:umap)
        # # Errors:
        # ids = findall(vec(round.(probs(M, data.X)) .!= data.y))
        # x_wrongly_labelled = data.X[:,ids]
        # scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
        plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
    end
end
plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
for (name, plts) in plot_dict
    plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
    savefig(plt, joinpath(www_path,"models_train_after_$(name)_strict.png"))
end
```

```{julia}
using Serialization
results = Serialization.deserialize(joinpath(output_path,"results_strict.jls"))
```

### 2022-07-31 dev/notebooks/synthetic.qmd
more work on gravitational generator. It works as expected if we choose strict convergence: we consider concergence as when the probability threshold is reached AND the reduction in our loss function from one iteration to another is small.
wuoohoo

```{julia}
n_evals = 5
n_rounds = 50
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 5
using Serialization
results = run_experiments(experiments;save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds, n_folds=n_folds)
Serialization.serialize(joinpath(output_path,"results.jls"),results)
```

@@

```{julia}
using Serialization
results = [Serialization.deserialize(joinpath(output_path,subfolder,"results.jls")) for subfolder in readdir(output_path)];
```

```{julia}
using Gadfly
plts = [plot(res) for res in values(results)]
```

>>>>

## Simple convergence

```{julia}
n_evals = 1
n_rounds = 1
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 1
n_bootstrap = 1
using Serialization
results = run_experiments(
    experiments;save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds,n_folds=n_folds,n_bootstrap=n_bootstrap
)
Serialization.serialize(joinpath(output_path,"results.jls"),results)
```

@@

## Strict convergence

```{julia}
n_evals = 1
n_rounds = 1
evaluate_every = Int(round(n_rounds/n_evals))
n_folds = 1
n_bootstrap = 1
using Serialization
results = run_experiments(
    experiments;
    save_path=output_path,evaluate_every=evaluate_every,n_rounds=n_rounds, n_folds=n_folds, n_bootstrap=n_bootstrap,
    convergence=:strict, save_name_suffix="strict"
)
Serialization.serialize(joinpath(output_path,"results_strict.jls"),results)
```

```{julia}
using AlgorithmicRecourseDynamics.Models: model_evaluation
plot_dict = Dict(key => Dict() for (key,val) in results)
fold = 1
for (name, res) in results
    exp_ = res.experiment
    plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
    rec_sys = exp_.recourse_systems[fold]
    sys_ids = collect(exp_.system_identifiers)
    M = length(rec_sys)
    for m in 1:M
        model_name, generator_name = sys_ids[m]
        M = rec_sys[m].model
        score = round(model_evaluation(M, exp_.test_data),digits=2)
        plt = plot(M, exp_.test_data, title="$name;\n $model_name ($score)")
        # Errors:
        ids = findall(vec(round.(probs(M, exp_.test_data.X)) .!= exp_.test_data.y))
        x_wrongly_labelled = exp_.test_data.X[:,ids]
        scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
        plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
    end
end
plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
for (name, plts) in plot_dict
    plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
    savefig(plt, joinpath(www_path,"models_test_after_$(name)_strict.png"))
end
```

```{julia}
using AlgorithmicRecourseDynamics.Models: model_evaluation
plot_dict = Dict(key => Dict() for (key,val) in results)
fold = 1
for (name, res) in results
    exp_ = res.experiment
    plot_dict[name] = Dict(key => [] for (key,val) in exp_.generators)
    rec_sys = exp_.recourse_systems[fold]
    sys_ids = collect(exp_.system_identifiers)
    M = length(rec_sys)
    for m in 1:M
        model_name, generator_name = sys_ids[m]
        M = rec_sys[m].model
        data = rec_sys[m].data
        score = round(model_evaluation(M, data),digits=2)
        plt = plot(M, data, title="$name;\n $model_name ($score)")
        # Errors:
        ids = findall(vec(round.(probs(M, data.X)) .!= data.y))
        x_wrongly_labelled = data.X[:,ids]
        scatter!(plt, x_wrongly_labelled[1,:], x_wrongly_labelled[2,:], ms=7.5, color=:red, label="")
        plot_dict[name][generator_name] = vcat(plot_dict[name][generator_name], plt)
    end
end
plot_dict = Dict(key => reduce(vcat, [plots[key] for plots in values(plot_dict)]) for (key, value) in generators)
for (name, plts) in plot_dict
    plt = plot(plts..., layout=(length(choices),length(models)),size=(length(choices)*300,length(models)*300))
    savefig(plt, joinpath(www_path,"models_train_after_$(name)_strict.png"))
end
```

```{julia}
using Serialization
results = Serialization.deserialize(joinpath(output_path,"results_strict.jls"))
```
